*Supplement* Hot and fresh: evidence of climate-related suboptimal conditions for seagrass in a large Gulf coast estuary

### 2024-05-08 manu-draft.qmd
more gam revisions, results done, working on discuss

load(file = here('data/speidat.RData'))
load(file = here('data/gamfit.RData'))
load(file = here('tabs/dattab.RData'))
load(file = here('tabs/temptrndtab.RData'))
load(file = here('tabs/saltrndtab.RData'))
load(file = here('tabs/mixdaytab.RData'))
load(file = here('tabs/modtab.RData'))
load(file = here('data/sgmods.RData'))
load(file = here('data/onb.RData'))
```

@@

For comparison to seagrass, the annual metrics calculated from the EPC data were referenced to approximate periods of time between the annual seagrass transect surveys, as opposed to the calendar year for describing trends above. Bay segment stressor metrics were calculated as the average counts in each "transect year" from all stations in each segment from 1998 to 2022. Preliminary analyses evaluated different lagged associations between the stressor metrics and seagrass change, although initial results suggested no additional insight could be gained using lagged assessments compared to the annual summaries. As such, the annual stressor metrics were compared to frequency occurrence (all species) each year by bay segment. GAMs were used to evaluate frequency occurrence in response to the independent variables, where the latter were the stressor metrics for temperature, salinity, or both. Additional predictors included year and light attenuation as estimated from Secchi depth [@tbep0696]. For 1998 to 2022, `r onb$epc` of the Secchi observations were recorded on the bottom. A single smooth term was used for each predictor using a thin plate regression spline, including a tensor product interaction term with a cubic regression spline that evaluated the potential interacting effects of each predictor with year. Two models were evaluated, one with the temperature and salinity metrics together (@eq-epcmod1) and another with the both metric (@eq-epcmod2). All models excluded Lower Tampa Bay because of minimal seagrass change over time. Additionally, total seagrass frequency occurrence was used as a response variable as an aggregate measure of community change to potential stressors given the likelihood that species-specific changes may be more difficult to model and that the majority of seagrass loss in recent years was dominated by *H. wrightii* in the upper bay segments (Figure [-@fig-seagrasschg]b). The GAM equations in mgcv R package notation [@wood17] are below, where $s$ is an smoother function for a single predictor and $ti$ is the tensor product interaction between predictors. All GAMs were fit using restricted maximum likelihood evaluation [@wood11].

@@

$$
\begin{aligned}
Seagrass \sim s(year, by = bay \ segment) + s(light \ att, by = bay \ segment) + 
\\
s(both, by = bay \ segment) +
\\
ti(light \ att, \ year, by, bay \ segment) + ti(both, year, by = bay \ segment)
\end{aligned}
$$ {#eq-epcmod2}
Separate models were constructed to provide an additional weight-of-evidence for the FIM and PDEM temperature and salinity datasets relative to seagrass change. These models used direct measurements of salinity and temperature as independent variables because the stressor metrics could not be calculated using the sampling designs from these monitoring programs (i.e., each sample was a distinct location). GAMs were used for the FIM data to evaluate seagrass percent cover using year, temperature, and salinity as independent variables. Separate smoothed terms for each predictor and iteractions with year for each predictor for each bay segment were included as above. GAMs for the PDEM data were constructed similarly, except only Old Tampa Bay was evaluated due to spatial limitations of the data. Neither the FIM nor PDEM models used light attenuation as a predictor variable given that a large percentage of Secchi observations were measured on the bottom (`r onb$fim` and `r onb$pinco`, respectively), providing further support that light environments have not been limiting for seagrasses. Model input data were further subset to include only months from July to October to describe seagrasses during the growing season and to reduce potential seasonal effects. Lastly, all data were averaged annually for the monthly subsets for comparability of sample size (i.e., power) with the EPC models. For the PDEM models, presence/absence was converted to frequency occurrence as the number of sites with seagrass in a year divided by the total number of sites and all data were subset to less than 2 meters to better characterize locations where seagrass occurs. The GAM equations for the FIM (@eq-fimmod) and PDEM (@eq-pincomod) datasets were as follows.

@@

```{r}
# epcmodsum1 <- modtxt_fun(sgmods$epcmod1)
# epcmodsum2 <- modtxt_fun(sgmods$epcmod2)
# fimmodsum <- modtxt_fun(sgmods$fimmod)
# secchi on bottom is na issue
# also averaging all models by year in the bay segment for comparable sample size, fim/pdem limited to july - nov to minimize seasonal effect and when transect surveys completed for comparability
# < 17% for epc, much higher for fim/pdem so la not included, but can show la relatively stable in a plot
# epc is bonk becuase of disconnect between sites
# fim is good, makes sense, but otb is weird
# pdem temp makes sense for otb
```

GAMs to assess the potential effects of year, light attenuation, temperature and salinity on seagrass change provided modest evidence that seagrass change was influenced by climate stressors. The model for the EPC data included the temperature and salinity metrics as potentially important for explaining inter-annual variation in seagrass frequency occurrence (). An interaction of the metrics with time period was also observed, with notable differences by predictor (Figure [-@fig-sgmod]a, @tbl-modtab). For the period prior to 2016, frequency occurrence of seagrass increased as the number of days where temperature was above 30$^\circ$ increased, whereas no relationship was observed after 2016. However, the interaction between time period and temperature was not significant, suggesting no difference between time periods. The opposite trend was observed for the number of days when salinity was below 25 ppt, where no relationship was observed prior to 2016 and a decreasing relationship was observed after 2016 (interaction term between salinity and time period was marginally significant, p \< 0.1, Figure [-@fig-sgmod]a, @tbl-modtab). The interaction between the temperature and salinity metric was included in the most parsimonious model, although it was not significant. The second model for the EPC data that evaluated the number of days when both temperature was above and salinity was below the threshold was also significant (), although the interaction term with time period was not included in the most parsimonious model, suggesting no effect of this metric in describing seagrass change after 2016 (@tbl-modtab). Note that [@fig-sgmod] shows model predictions for OTB only, although the shape of the relationships do not change for other bay segments other than as additive shifts of the curves up or down.

The most parsimonious model for the FIM dataset also showed potential temperature and salinity associations with seagrass change after 2016, although explanatory power was lower than the EPC models likely due to larger spread of the data (). Increases in temperature and decreases in salinity were both associated with reductions in percent cover after 2016, whereas no relationships were observed prior to 2016 (p \< 0.05 and p \< 0.1 for the interactions between temperature or salinity with time period, respectively, Figure [-@fig-sgmod]b, @tbl-modtab). The interaction between salinity and temperature was not included as a significant term in the final FIM model. Lastly, the most parsimonious model for the PDEM dataset only included time period as a marginally significant variable and the overall model fit was not significant (@tbl-modtab).

@@

Global increases in temperature and altered precipitation patterns related to climate change have had measurable effects on the structure and functioning of a wide range of natural environments [@osland2015; @oliver2018]. For Tampa Bay, these changes have been demonstrated using long-term trends in water temperature and salinity, which mirrored long-term changes in air temperature and precipitation. Tampa Bay has gotten hotter and fresher; water temperature has increased by 0.03 - 0.04 $^\circ$C per year and salinity has decreased by 0.04 - 0.06 ppt per year, translating to an increase of 1.3 to 1.7 $^\circ$C and a decrease of 1.6 to 2.6 ppt over the past fifty years. These changes were demonstrated in three long-term datasets with different sampling methods and periods of record. Understandably, the trends were most clearly observed in the dataset with the longest period of record (EPC), covering nearly fifty years of monthly observations. These long-term changes manifested into consistent trends in known seagrass stressors; the continuous number of days increased each year when temperature, salinity, or both crossed thresholds.

Similar regional, long-term changes in coastal waters and estuaries have been observed by others [@carlson2018; @nickerson2023; @shiir]. @nickerson2023 evaluated sea surface trends at a larger spatial scale for Tampa Bay, the West Florida Continental Shelf, and the adjacent Gulf of Mexico. Temperature trends were similar to those herein for Tampa Bay (the EPC dataset was also used). @nickerson2023 also noted that temperature increases in Tampa Bay were most pronounced in the winter, although they rightfully acknowledge the sensitivity of their results to conditions at the start and end of the time series. Our assessment evaluated non-parametric trends (i.e., Kendall tests, less sensitive to outliers) at individual EPC stations and bay segments, whereas @nickerson2023 evaluated the EPC data as an average for the entire bay for consistency of comparison to their larger spatial area and model domain. Our results showing increases in temperature and decreases in salinity in the summer, early fall, most notably for OTB and northern stations of HB [@fig-kendall], are likely related to hydrodynamic characteristics of these segments relative to MTB and LTB that flush more regularly with the Gulf of Mexico. These upper bay segments are more affected by hydrologic inflows (HB), lack of circulation (OTB), or thermal stress related to more rapid warming with shallower depths. The significant reduction in salinity for LTB is also of note, perhaps related to gravitational circulation patterns that export lower salinity water from upstream in the main shipping channels [@weisberg2006]. Additionally, @shiir provided a recent assessment of a 2023 heatwave in south Florida, supported by a 20-year trend assessment that suggested estuaries were warming at nearly double the rate of the Gulf of Mexico. The upper limit of our warming estimate for Tampa Bay is comparable. Notably, @carlson2018 suggest a link between historical seagrass losses in Florida Bay and rapid warming in shallow areas with low surface reflectance.

Our relatively simple modeling approach provided some evidence that climate-related stressors impart some effect on recent seagrass losses in Tampa Bay. The models did not provide a consistent, nor statistically powerful, explanation that increasing temperature and decreasing salinity were key (or the sole) drivers. However, evaluating all models together as weight-of-evidence suggests there is value in considering multiple datasets and models to interpret noisy patterns and compounding ecological processes. Model results for the EPC and FIM datasets both suggested that increasing temperature and decreasing salinity were associated with potential seagrass loss post 2016, described primarily using separate interaction terms of temperature or salinity with time period. For the EPC model, the interaction term was not significant for temperature, whereas the interaction was marginally significant for the salinity metric, such that a negative association between seagrass and salinity stress was observed post 2016 as compared to pre 2016. Likewise, the FIM models had a significant interaction term for the association of temperature with time period and a marginally significant interaction term for the association of salinity with time period. An important distinction between the EPC and FIM models is that the former evaluated the number of days above/below thresholds each year to quantify annual temperature or salinity stress, whereas the latter evaluated observed temperature and salinity values at the time of seagrass sampling. As such, both models attempted to describe the role of these stressors on potential seagrass change, but use different independent variables given the different sampling designs of each monitoring program. These differences highlight challenges describing autecological relationships in long-term datasets, while also demonstrating the utility of our weight-of-evidence approach to describe such relationships.

@@

![Predicted results from Generalized Additive Models evaluating seagrass changes versus year, temperature, and salinity for the (a) FIM (first three columns evaluating mean annual percent cover from 0-100, n = 81, R.sq. 0.81, deviance explained 89%) and (b) PDEM (right column evaluating annual frequency occurrence from 0-1 in OTB only, n = 19, R.sq. 0.69, deviance explained 81%) datasets. Partial effects describe the modeled association between each predictor and seagrass change after accounting for the effects of the other predictors. The models also included smoothed interaction terms with year (not shown). See Tables S9 and S10 for additional model fit statistics. OTB: Old Tampa Bay, HB: Hillsborough Bay, MTB: Middle Tampa Bay.](figs/sgmods.png){#fig-sgmod}

@@

```{r}
#| label: tbl-modtab
#| tbl-cap: 'Summary of the most parsimonious Generalized Linear Models (GLMs) developed to evaluate the effects of temperature and salinity variables on seagrass change between time periods. Models were developed for different datasets described in the model column. The significant intercept term for each model is not shown and described factor conditions not explicitly shown by the remaining terms (e.g., OTB and pre-2016). The model term estimates include the standard error in parentheses. Note that the "Temp" and "Sal" terms for the EPC models describe metrics that measure the number of days when temperature was above or salinity was below a threshold value, whereas the terms in all other models describe observed temperature or salinity (see text for description). ** p < 0.005, * p < 0.05, $^.$ p < 0.1. EPC: Environmental Protection Commission, FIM: Fisheries Independent Monitoring, PDEM: Pinellas County Department of Environmental Management. OTB: Old Tampa Bay, HB: Hillsborough Bay, MTB: Middle Tampa Bay.'
modtab
```

>>>>

load(file = here('data/speidat.RData'))
load(file = here('data/gamfit.RData'))
load(file = here('tabs/dattab.RData'))
load(file = here('tabs/temptrndtab.RData'))
load(file = here('tabs/saltrndtab.RData'))
load(file = here('tabs/mixdaytab.RData'))
load(file = here('data/sgmods.RData'))
load(file = here('data/sgmodsum.RData'))
load(file = here('data/onb.RData'))
```

@@

For comparison to seagrass, the annual metrics calculated from the EPC data were referenced to approximate periods of time between the annual seagrass transect surveys, as opposed to the calendar year for describing trends above. Bay segment stressor metrics were calculated as the average counts in each "transect year" from all stations in each segment from 1998 to 2022. Preliminary analyses evaluated different lagged associations between the stressor metrics and seagrass change, although initial results suggested no additional insight could be gained using lagged assessments compared to the transect year summaries. As such, the stressor metrics were compared to frequency occurrence (all species) each year by bay segment. GAMs were used to evaluate frequency occurrence in response to the independent variables, where the latter were the stressor metrics for temperature, salinity, or both. Additional predictors included year and light attenuation as estimated from Secchi depth [@tbep0696]. For 1998 to 2022, `r onb$epc` of the Secchi observations were recorded on the bottom. A single smooth term was used for each predictor using a thin plate regression spline, including a tensor product interaction term with a cubic regression spline that evaluated the potential interacting effects of each predictor with year. Two models were evaluated, one with the temperature and salinity metrics together (@eq-epcmod1) and another with the both metric (@eq-epcmod2). All models excluded Lower Tampa Bay because of minimal seagrass change over time. Additionally, total seagrass frequency occurrence was used as a response variable as an aggregate measure of community change to potential stressors given the likelihood that species-specific changes may be more difficult to model and that the majority of seagrass loss in recent years was dominated by *H. wrightii* in the upper bay segments (Figure [-@fig-seagrasschg]b). The GAM equations in mgcv R package notation [@wood17] are below, where $s$ is a smoother function for a single predictor and $ti$ is the tensor product interaction between predictors. All GAMs were fit using restricted maximum likelihood evaluation [@wood11].

@@

$$
\begin{aligned}
Seagrass \sim s(year, by = bay \ segment) + s(light \ att, by = bay \ segment) + 
\\
s(both, by = bay \ segment) +
\\
ti(light \ att, \ year, by, bay \ segment) + ti(both, year, by = bay \ segment)
\end{aligned}
$$ {#eq-epcmod2} models were constructed to provide an additional weight-of-evidence for the FIM and PDEM temperature and salinity datasets relative to seagrass change. These models used direct measurements of salinity and temperature as independent variables because the stressor metrics could not be calculated using the sampling designs from these monitoring programs (i.e., each sample was a distinct location). GAMs were used for the FIM data to evaluate annual average seagrass percent cover using year, temperature, and salinity as independent variables. Separate smoothed terms for each predictor and interactions with year for each predictor for each bay segment were included as above. GAMs for the PDEM data were constructed similarly, except only Old Tampa Bay was evaluated due to spatial limitations of the data. Neither the FIM nor PDEM models used light attenuation as a predictor variable given that a large percentage of Secchi observations were measured on the bottom (`r onb$fim` and `r onb$pinco`, respectively), providing further support that light environments have not been limiting for seagrasses. Model input data were further subset to include only months from July to October to describe seagrasses during the growing season and to reduce potential seasonal effects. Lastly, all data were averaged annually for the monthly subsets for comparability of sample size (i.e., power) with the EPC models. For the PDEM models, presence/absence was converted to frequency occurrence as the number of sites with seagrass in a year divided by the total number of sites and all data were subset to less than 2 meters to better characterize locations where seagrass occurs. The GAM equations for the FIM (@eq-fimmod) and PDEM (@eq-pincomod) datasets, respectively, were as follows.

@@

```{r}
sgmodtxt <- modtxt_fun(sgmodsum)
```

GAMs to assess the potential effects of year, light attenuation, temperature and salinity on seagrass change provided some evidence that seagrass change was influenced by climate stressors, particularly for the FIM and PDEM datasets. The model for the EPC data (`r sgmodtxt$epcmod1`) had notable associations with year for the OTB and MTB bay segments and for light attenuation for the HB bay segment (@fig-sgmod1, Table S7). Associations of seagrass change with the temperature and salinity metrics were weak and somewhat contrary to expectation (i.e., an increasing partial effect on seagrass as the stressor metrics increased). The EPC model using the "both" stressor metric was similar (`r sgmodtxt$epcmod2`, Figure S7, Table S8), suggesting no added benefit of evaluating times when both temperature was above and salinity was below the threshold.

The GAMs for the FIM and PDEM datasets had notable associations with temperature and salinity stress, where an increase in temperature and a decrease in salinity was often associated with a decline in seagrass depending on bay segment. The FIM model (`r sgmodtxt$fimmod`), Figure \[-@fig-sgmod2\]a, Table S9) showed a slight decrease in seagrass with increasing temperature for HB and MTB, with a similar change as salinity decreased. The PDEM model (`r sgmodtxt$pincomod`, Figure \[-@fig-sgmod2\]b, Table S10) that evaluated only OTB showed a distinct decrease with seagrass as temperature increased, but no association with salinity. Some of the interaction terms of the predictors with year suggested the relationships with seagrass may have changed over time, but the results were not consistent between bay segments or models (Tables S7 to S10).

@@

Global increases in temperature and altered precipitation patterns related to climate change have had measurable effects on the structure and functioning of a wide range of natural environments [@osland2015; @oliver2018]. For Tampa Bay, these changes have been demonstrated using long-term trends in water temperature and salinity, which mirrored long-term changes in air temperature and precipitation. Tampa Bay has gotten hotter and fresher; water temperature has increased by 0.03 - 0.04 $^\circ$C per year and salinity has decreased by 0.04 - 0.06 ppt per year, translating to an increase of 1.3 to 1.7 $^\circ$C and a decrease of 1.6 to 2.6 ppt over the past fifty years. The number of days each year when temperature was above 30 $^\circ$C or salinity was below 25 ppt has also increased consistently, with an average increase baywide of 48 and 37 days, respectively, since 1975. These changes were demonstrated in three long-term datasets with different sampling methods and periods of record. Understandably, the trends were most clearly observed in the dataset with the longest period of record (EPC), covering nearly fifty years of monthly observations. These long-term changes manifested into consistent trends in known seagrass stressors; the continuous number of days increased each year when temperature, salinity, or both crossed thresholds.

Similar regional, long-term changes in coastal waters and estuaries have been observed by others [@carlson2018; @nickerson2023; @shiir]. @nickerson2023 evaluated sea surface trends at a larger spatial scale for Tampa Bay, the West Florida Continental Shelf, and the adjacent Gulf of Mexico. Temperature trends were similar to those herein for Tampa Bay (the EPC dataset was also used). @nickerson2023 also noted that temperature increases in Tampa Bay were most pronounced in the winter, although they rightfully acknowledge the sensitivity of their results to conditions at the start and end of the time series. Our assessment evaluated non-parametric trends (i.e., Kendall tests, less sensitive to outliers) at individual EPC stations and bay segments, whereas @nickerson2023 evaluated the EPC data as an average for the entire bay for consistency of comparison to their larger spatial area and model domain. Our results showing increases in temperature and decreases in salinity in the summer, early fall, most notably for OTB and northern stations of HB (@fig-kendall), are likely related to hydrodynamic characteristics of these segments relative to MTB and LTB that flush more regularly with the Gulf of Mexico. These upper bay segments are more affected by hydrologic inflows (HB), lack of circulation (OTB), or thermal stress related to more rapid warming with shallower depths. The significant reduction in salinity for LTB is also of note, perhaps related to gravitational circulation patterns that export lower salinity water from upstream in the main shipping channels \[@weisberg2006\]. Additionally, @shiir provided a recent assessment of a 2023 heatwave in south Florida, supported by a 20-year trend assessment that suggested estuaries were warming at nearly double the rate of the Gulf of Mexico. The upper limit of our warming estimate for Tampa Bay is comparable. Notably, @carlson2018 suggest a link between historical seagrass losses in Florida Bay and rapid warming in shallow areas with low surface reflectance.

Our relatively simple modeling approach provided some evidence that climate-related stressors impart some effect on recent seagrass losses in Tampa Bay. The models did not provide a consistent explanation that increasing temperature and decreasing salinity were key (or the sole) drivers. However, evaluating all models together as weight-of-evidence suggests there is value in considering multiple datasets and models to interpret noisy patterns and compounding ecological processes. Model results for the FIM and PDEM datasets both suggested that increasing temperature and decreasing salinity were associated with potential seagrass loss in recent years (@fig-sgmod2). For the EPC model, none of the stressor metrics were associated with seagrass change. An important distinction between the EPC and other models is that the former evaluated the number of days above/below thresholds each year to quantify annual temperature or salinity stress, whereas the latter evaluated observed temperature and salinity values at the time of seagrass sampling. As such, both models attempted to describe the role of these stressors on potential seagrass change, but use different independent variables given the different sampling designs of each monitoring program. These differences highlight challenges describing autecological relationships in long-term datasets, while also demonstrating the utility of our weight-of-evidence approach to describe such relationships.

@@

![Partial effects of smoothers (rows) by bay segment (columns) from a Generalized Additive Model used to describe seagrass change relative to year, light attenuation, the number of days each year when temperature was above 30 $^\circ$C, and the number of days each year when salinity was below 25 ppt. The EPC data were used for the independent variables in the model. Partial effects describe the modeled association between each predictor and seagrass frequency occurrence after accounting for the effects of the other predictors. The model also included smoothed interaction terms with year (not shown). See Table S7 for additional model fit statistics. OTB: Old Tampa Bay, HB: Hillsborough Bay, MTB: Middle Tampa Bay. `r modtxt_fun(sgmodsum)$epcmod1`](figs/sgmod1.png){#fig-sgmod1}

![Partial effects of smoothers (rows) by bay segment (columns) from Generalized Additive Models evaluating seagrass changes versus year, temperature, and salinity for the (a) FIM (first three columns evaluating mean annual percent cover from 0-100, `r modtxt_fun(sgmodsum)$fimmod`) and (b) PDEM (right column evaluating annual frequency occurrence from 0-1 in OTB only, `r modtxt_fun(sgmodsum)$pincomod`) datasets. Partial effects describe the modeled association between each predictor and seagrass change after accounting for the effects of the other predictors. The models also included smoothed interaction terms with year (not shown). See Tables S9 and S10 for additional model fit statistics. OTB: Old Tampa Bay, HB: Hillsborough Bay, MTB: Middle Tampa Bay.](figs/sgmod2.png){#fig-sgmod2}

@@

### 2024-05-08 supplement.qmd
more gam revisions, results done, working on discuss

load(file = here('tabs/supptempfittab.RData'))
load(file = here('tabs/suppsalifittab.RData'))
load(file = here('tabs/supp1mixtab.RData'))
load(file = here('tabs/supp2mixtab.RData'))
load(file = here('tabs/supp1daytab.RData'))
load(file = here('tabs/supp2daytab.RData'))
load(file = here('tabs/suppepcmod1tab.RData'))
load(file = here('tabs/suppepcmod2tab.RData'))
load(file = here('tabs/suppfimmodtab.RData'))
load(file = here('tabs/supppincomodtab.RData'))
```

@@

![Partial effects of smoothers by bay segment from a Generalized Additive model used to describe seagrass change relative to year, light attenuation, the number of days each year when temperature was above 30 $^\circ$C, and the number of days each year when salinity was below 25 ppt. The EPC data were used to develop the model. Partial effects describe the modeled association between each predictor and seagrass frequency occurrence after accounting for the effects of the other predictors. The model also included smoothed interaction terms with year (not shown). See Table S7 for additional model fit statistics. OTB: Old Tampa Bay, HB: Hillsborough Bay, MTB: Middle Tampa Bay. n = 75, R.sq. 0.93, Deviance explained 97\%](figs/suppepcmod1.png){#fig-suppepcmod1}
{{< pagebreak >}}

![Partial effects of smoothers by bay segment from a Generalized Additive model used to describe seagrass change relative to year, light attenuation, and the number of days each year when both temperature was above 30 $^\circ$C and salinity was below 25 ppt. The EPC data were used to develop the model. Partial effects describe the modeled association between each predictor and seagrass frequency occurrence after accounting for the effects of the other predictors. The model also included smoothed interaction terms with year (not shown). See Table S8 for additional model fit statistics. OTB: Old Tampa Bay, HB: Hillsborough Bay, MTB: Middle Tampa Bay. n = 75, R.sq. 0.93, Deviance explained 97%](figs/suppepcmod2.png){#fig-suppepcmod2}
{{< pagebreak >}}

>>>>

source(here('R/funcs.R'))

load(file = here('data/sgmodsum.RData'))
load(file = here('tabs/supptempfittab.RData'))
load(file = here('tabs/suppsalifittab.RData'))
load(file = here('tabs/supp1mixtab.RData'))
load(file = here('tabs/supp2mixtab.RData'))
load(file = here('tabs/supp1daytab.RData'))
load(file = here('tabs/supp2daytab.RData'))
load(file = here('tabs/suppepcmod1tab.RData'))
load(file = here('tabs/suppepcmod2tab.RData'))
load(file = here('tabs/suppfimmodtab.RData'))
load(file = here('tabs/supppincomodtab.RData'))
```

@@

![Partial effects of smoothers (rows) by bay segment (columns) from a Generalized Additive model used to describe seagrass change relative to year, light attenuation, and the number of days each year when both temperature was above 30 $^\circ$C and salinity was below 25 ppt. The EPC data were used to develop the model. Partial effects describe the modeled association between each predictor and seagrass frequency occurrence after accounting for the effects of the other predictors. The model also included smoothed interaction terms with year (not shown). See Table S8 for additional model fit statistics. OTB: Old Tampa Bay, HB: Hillsborough Bay, MTB: Middle Tampa Bay. `r modtxt_fun(sgmodsum)$epcmod2`](figs/suppsgmod3.png){#fig-suppsgmod3}
